### 2022-11-05 dissertation/chapters/05-probability.tex
🚧 (chapter6): Save point

\noindent
Probability theory and statistics can be traced back to as early as the 18th century. In 1718,~\citeauthor{ref:demoivre:1718}~\cite{ref:demoivre:1718} published the \textit{Doctrine of of Chance}; a book that is widely seen as the first published book on probability theory. In 1763, Thomas Bayes~\cite{ref:bayes:1763} published an article titled \textit{An Essay towards solving a Problem in the Doctrine of Chances} where the first version of Bayes' Theorem was introduced.

@@

      \item \textbf{Section \ref{sec:probability:bayes_theorem}} introduces Bayes' Theorem, the fundamental theorem upon which the \acs{BHH} is built.

@@

\section{Bayes' Theorem}\label{sec:probability:bayes_theorem}

Bayes' Theorem, named after Thomas Bayes, describes the probability of an event $A$, based on prior knowledge of conditions that might be related to $A$~\cite{ref:zalta:2015}. In order to derive the formal theorem and proof, first consider Definition \ref{def:probability:bayes_theorem:partition} below~\cite{ref:zalta:2015}.
\\
\begin{definition}
      \label{def:probability:bayes_theorem:partition}
      For some positive integer, $K$, let the sets $B_{1}, B_{2}, \dots, B_{K}$ be such that

@@

\begin{theorem}[\textbf{Bayes' Theorem}]
      \label{th:probability:bayes_theorem:theorem}
      Assume that $[B_{1}, B_{2}, \dots, B_{K}]$ is a partition of $S$ such that $P(B_{i}) > 0$, for $i = 1,2, \dots, K$ then\\
      \\
      \begin{equation}
            P(B_{j} \vert A) = \frac{P(A \vert B_{j})P(B_{j})}{\sum_{i=1}^{K} P(A \vert B_{i})P(B_{i})}
      \end{equation}
\end{theorem}

@@

\noindent
One of the many applications of Bayes' Theorem is to do statistical inference. Bayes' Theorem expresses how a degree of belief, expressed as a probability, should rationally change to account for the availability of related evidence.

@@

\noindent
By Bayes' Theorem, the posterior distribution with given prior data $\mathcal{D}$ is given in Equation~\eqref{eq:probability:conjugate_priors:binom_likelihood:posterior} below.

@@

\noindent
By Bayes' Theorem, the posterior distribution with given prior data $\mathcal{D}$ is given in Equation~\eqref{eq:probability:conjugate_priors:mult_likelihood:posterior} below.

@@

In general, there are two main views to probability and statistics. These include the \textit{frequentist} and the \textit{Bayesian} view of statistics. Naturally, Bayesian statistics is based on Bayes' Theorem as was presented in Section \ref{sec:probability:bayes_theorem}. Bayesian statistics describe the probability of an event in terms of some belief, based on previous knowledge of the event, and the conditions under which the event happens~\cite{ref:hackenberger:2019}. To introduce the concept of Bayesian inference and analysis, one first need to understand the difference between these two approaches. Bayesian statistics out-date the frequentist approach, but lacked interest in the early days, partly because of the limited applications where the conjugate priors where known~\cite{ref:hackenberger:2019}. More recent advancements in mathematical methods popularised the Bayesian approach again. A notable contribution to this switch, was the development of \acf{MCMC} in the 1950's. This family of algorithms allowed for the construction of random sampling algorithms from a probability distribution, which allows for the calculation of Bayesian hierarchical models~\cite{ref:hackenberger:2019}. Soon after followed one of the earliest papers that use Bayesian statistics in the field of medicine in 1982~\cite{ref:ashby:2006}.

@@

\noindent
Finally, the posterior density of $\theta \vert y_{1}, y_{2}, \dots, y_{n}$, according to Bayes' Theorem, is given in Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:posterior_density} below.

>>>>

\noindent
Probability theory and statistics can be traced back to as early as the 18th century. In 1718,~\citeauthor{ref:demoivre:1718}~\cite{ref:demoivre:1718} published the \textit{Doctrine of of Chance}; a book that is widely seen as the first published book on probability theory. In 1763, Thomas Bayes~\cite{ref:bayes:1763} published an article titled \textit{An Essay towards solving a Problem in the Doctrine of Chances} where the first version of \index{Bayes' Theorem}Bayes' Theorem was introduced.

@@

      \item \textbf{Section \ref{sec:probability:bayes_theorem}} introduces \index{Bayes' Theorem}Bayes' Theorem, the fundamental theorem upon which the \acs{BHH} is built.

@@

\index{Bayes Theorem}
\section{Bayes' Theorem}\label{sec:probability:bayes_theorem}

\index{Bayes' Theorem}Bayes' Theorem, named after Thomas Bayes, describes the probability of an event $A$, based on prior knowledge of conditions that might be related to $A$~\cite{ref:zalta:2015}. In order to derive the formal theorem and proof, first consider Definition \ref{def:probability:bayes_theorem:partition} below~\cite{ref:zalta:2015}.
\\
\begin{definition}
      \label{def:probability:bayes_theorem:partition}
      For some positive integer, $K$, let the sets $B_{1}, B_{2}, \dots, B_{K}$ be such that

@@

\index{Bayes Theorem}
\begin{theorem}[\textbf{Bayes' Theorem}]
      \label{th:probability:bayes_theorem:theorem}
      Assume that $[B_{1}, B_{2}, \dots, B_{K}]$ is a partition of $S$ such that $P(B_{i}) > 0$, for $i = 1,2, \dots, K$ then\\
      \\
      \begin{equation}
            P(B_{j} \vert A) = \frac{P(A \vert B_{j})P(B_{j})}{\sum_{i=1}^{K} P(A \vert B_{i})P(B_{i})}
      \end{equation}
\end{theorem}

@@

\noindent
One of the many applications of \index{Bayes' Theorem}Bayes' Theorem is to do statistical inference. \index{Bayes' Theorem}Bayes' Theorem expresses how a degree of belief, expressed as a probability, should rationally change to account for the availability of related evidence.

@@

\noindent
By \index{Bayes' Theorem}Bayes' Theorem, the posterior distribution with given prior data $\mathcal{D}$ is given in Equation~\eqref{eq:probability:conjugate_priors:binom_likelihood:posterior} below.

@@

\noindent
By \index{Bayes' Theorem}Bayes' Theorem, the posterior distribution with given prior data $\mathcal{D}$ is given in Equation~\eqref{eq:probability:conjugate_priors:mult_likelihood:posterior} below.

@@

In general, there are two main views to probability and statistics. These include the \textit{frequentist} and the \textit{Bayesian} view of statistics. Naturally, Bayesian statistics is based on \index{Bayes' Theorem}Bayes' Theorem as was presented in Section \ref{sec:probability:bayes_theorem}. Bayesian statistics describe the probability of an event in terms of some belief, based on previous knowledge of the event, and the conditions under which the event happens~\cite{ref:hackenberger:2019}. To introduce the concept of Bayesian inference and analysis, one first need to understand the difference between these two approaches. Bayesian statistics out-date the frequentist approach, but lacked interest in the early days, partly because of the limited applications where the conjugate priors where known~\cite{ref:hackenberger:2019}. More recent advancements in mathematical methods popularised the Bayesian approach again. A notable contribution to this switch, was the development of \acf{MCMC} in the 1950's. This family of algorithms allowed for the construction of random sampling algorithms from a probability distribution, which allows for the calculation of Bayesian hierarchical models~\cite{ref:hackenberger:2019}. Soon after followed one of the earliest papers that use Bayesian statistics in the field of medicine in 1982~\cite{ref:ashby:2006}.

@@

\noindent
Finally, the posterior density of $\theta \vert y_{1}, y_{2}, \dots, y_{n}$, according to \index{Bayes' Theorem}Bayes' Theorem, is given in Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:posterior_density} below.
